Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-02-12 paper.tex
A significant contribution from Marina Karsanina

The large portion of the research papers mentioned above
\cite{rozenbaum2014,karsanina2015,ledesma2018,derossi2019,portillo2018developing,
  takada2003three,hopkins2013stars,jiao2013,Torquato_book,roding2020predicting,
  KarsaninaEJSS,capek2009,gerke2019tensor,REVpaper,LavrukhinPRE,Adler_recon,Y-T,
  tahmasebiPRL,Euras2012,EPL2,karsaninaPRL,jiao2007,SciRep1,Havelka,Geoderma2018}
is based on so-called correlation functions (CFs). These functions define a
probability that some events are occurring withing the structure (with some
special cases of surface functions), for example, 2-point probability $S_2$
measures the probability that both ends of the line segment are within the same
phase, 2-point cluster function $C_2$ -- that both ends of the line segment are
within the same interconnected cluster of same phase, and lineal path $L_2$
measures the probability that the whole segment lies within the phase (see
\highlight{Fig.1}). Compared to other descriptors, CFs possess a number of
features making them, in our opinion, a perfect candidate for universal
descriptor:
\begin{enumerate}
  \item They contain the majority of other metrics inside. For example, porosity
    is $S_2(0)$ (and at zero correlation length of the majority of correlation
    functions except for surface ones), surface area is represented by in the
    derivative at $S_2(0)$ \cite{debye1957scattering}. Connectivity is described by cluster
    functions $C_2$ or $C_3$. Pore-size distribution is related to pore-size
    function and partially by $L_2$ and $S_3$. Surface-void functions are
    interconnected with mean curvature and Euler characteristic
    \cite{ma2020generation}. Altogether, the classic set of correlation
    functions describes both the geometry and topology of the structure at hand;
  \item While the extent to which such classical set describes any arbitrary
    structure is not yet known, compared to other complex descriptors where are
    known pathways to assess the information content of correlation functions --
    there is a methodology for $S_2$ \cite{Gommes2} and approaches for other
    functions \cite{Degeneraty.045306,cherkasov2023towards};
  \item In addition to a possibility to compute correlation functions using
    images or other structural data, i.e., similar to all other descriptors,
    some CFs can be also determined experimentally
    \cite{debye1957scattering,barrall1992nmr,dietrich1995scattering,li2018accurate};
  \item Correlation functions can be scaled \cite{karsaninaPRL} or manipulated,
    for example, to add functions into a single one \cite{moctezuma2002};
  \item As opposed to scalar metrics such as porosity, surface area or Minkowski
    functionals, CFs can be computed in directions and describe anisotropic
    structures;
  \item Numerous rigorous bounds to obtain thermal, elastic, electromagnetic and
    transport properties are known, among which the most known (and least useful
    due to the impractically huge gap between the bounds) are Hashin-Strikman
    bounds \cite{hashin1963variational}. Other expressions based on integration
    of higher order $S_n$ are known and were found to be close to direct
    simulations
    \cite{brown1955solid,beran1965use,milton1981bounds,hlushkou2015effective};
  \item One can control both the size of the CFs-based descriptor and the
    information content via correlation length, number of functions in the set
    (including directions \cite{EPL2}), and by the order of functions (e.g., 2-
    and 3-point versions, or even more points).
\end{enumerate}
This list highlights that correlations functions are indeed possess necessary
properties so satisfy key criteria stated above.

>>>>

\begin{figure*}[ht]
  \centering
  \includegraphics[width=0.8\linewidth]{images/3point.png}
  \caption[]{A schematic representation of a binary porous media (pores are
    shaded) with examples of CFs computations (all classic two- and three-point
    correlation functions are presented according to the legend on the left
    side). Computations take into account all possible configurations of a line
    segment which either make a contribution to the CF (conventionally shown as
    ``CF = 1'') or do not make a contribution (``CF = 0'').}
  \label{fig:3point-scheme}
\end{figure*}
The large portion of the research papers mentioned above
\cite{rozenbaum2014,karsanina2015,ledesma2018,derossi2019,portillo2018developing,
  takada2003three,hopkins2013stars,jiao2013,Torquato_book,roding2020predicting,
  KarsaninaEJSS,capek2009,gerke2019tensor,REVpaper,LavrukhinPRE,Adler_recon,Y-T,
  tahmasebiPRL,Euras2012,EPL2,karsaninaPRL,jiao2007,SciRep1,Havelka,Geoderma2018}
is based on so-called correlation functions (CFs). These functions define a
probability that some events are occurring withing the structure (with some
special cases of surface functions), for example, 2-point probability $S_2$
measures the probability that both ends of the line segment are within the same
phase, 2-point cluster function $C_2$ -- that both ends of the line segment are
within the same interconnected cluster of same phase, and lineal path $L_2$
measures the probability that the whole segment lies within the phase (see
\cref{fig:3point-scheme}). Compared to other descriptors, CFs possess a number
of features making them, in our opinion, a perfect candidate for universal
descriptor:
\begin{enumerate}
  \item They contain the majority of other metrics inside. For example, porosity
    is $S_2(0)$ (and at zero correlation length of the majority of correlation
    functions except for surface ones), surface area is represented by in the
    derivative at $S_2(0)$ \cite{debye1957scattering}. Connectivity is described by cluster
    functions $C_2$ or $C_3$. Pore-size distribution is related to pore-size
    function and partially by $L_2$ and $S_3$. Surface-void functions are
    interconnected with mean curvature and Euler characteristic
    \cite{ma2020generation}. Altogether, the classic set of correlation
    functions describes both the geometry and topology of the structure at hand;
  \item While the extent to which such classical set describes any arbitrary
    structure is not yet known, compared to other complex descriptors where are
    known pathways to assess the information content of correlation functions --
    there is a methodology for $S_2$ \cite{Gommes2} and approaches for other
    functions \cite{Degeneraty.045306,cherkasov2023towards};
  \item In addition to a possibility to compute correlation functions using
    images or other structural data, i.e., similar to all other descriptors,
    some CFs can be also determined experimentally
    \cite{debye1957scattering,barrall1992nmr,dietrich1995scattering,li2018accurate};
  \item Correlation functions can be scaled \cite{karsaninaPRL} or manipulated,
    for example, to add functions into a single one \cite{moctezuma2002};
  \item As opposed to scalar metrics such as porosity, surface area or Minkowski
    functionals, CFs can be computed in directions and describe anisotropic
    structures;
  \item Numerous rigorous bounds to obtain thermal, elastic, electromagnetic and
    transport properties are known, among which the most known (and least useful
    due to the impractically huge gap between the bounds) are Hashin-Strikman
    bounds \cite{hashin1963variational}. Other expressions based on integration
    of higher order $S_n$ are known and were found to be close to direct
    simulations
    \cite{brown1955solid,beran1965use,milton1981bounds,hlushkou2015effective};
  \item One can control both the size of the CFs-based descriptor and the
    information content via correlation length, number of functions in the set
    (including directions \cite{EPL2}), and by the order of functions (e.g., 2-
    and 3-point versions, or even more points).
\end{enumerate}
This list highlights that correlations functions are indeed possess necessary
properties so satisfy key criteria stated above.
